Floral selectivity in mountain bumble bees and a test of the 'target use hypothesis

### 2021-06-15 ms_2/ms_2.Rmd
almost there...

Sampling of bumble- bee-wildflower interactions consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Sampling was only conducted during periods without rain and with an air temperature of at least 6°C. Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors. Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz. While male bumble bees were recorded, we chose to analyze only the pooled visitation of queens and workers. In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro. While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012. The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July. In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient. See Appendix X for a full description of sampling patterns.

@@

We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of selectivity between bumble bee species and floral morphotypes. We used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values. Then, using the results of this analysis, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf]. Finally, we compared morphotype selectivity across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype selectivity matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.

@@

@@

@@

It is well-known that bumble bees forage non-randomly accordingly to morphological constraints, innate preferences, and experienced rewards [e.g. Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm]. In this regard, the results of our null model analysis are as unsurprising as they are unambiguous; bumble bee resource selection almost never represents flora in unaltered proportion to their abundance. The *target use hypothesis* [@De_Valpine2013-be; @Harmon-Threatt2017-bz], however, extends this idea to the perfect antithesis of the random foraging model, positing that bumble bee foraging is a process of *pure* preference, completely irrespective of resource abundance.

Given the well-established mechanistic link between bumble bee tongue length and floral selection [e.g. @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained selectivity (i.e. null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice. This contradicts the *target use hypothesis* as a general rule of bumble bee foraging. It does not, however, mean that the *target use hypothesis* does not obtain under certain circumstances, such as when a highly preferred resource varies within a nonlimiting rage of availability. Moreover, our results suggest something akin to the *target use hypothesis* but formulated negatively with respect to resources that are *not selected*. Visitation rates to disc flowers in our system were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not. In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to parse these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging. Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study.

## Patterns of selectivity

We found that positive selection was focused on 3 floral morphotypes: bell, lip, and flag flowers. These results are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il in the Austrian Alps ~55 km south of our study system, suggesting that our findings are at least regionally generalizeable. When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (especially Syrphid hoverflies) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g. disc flowers) that receive less attention from bumble bees [@Neumayer1999-il]. In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper flower morphotypes. The process of floral resource partitioning *among* bumble bee species is nested within the broader process of floral partitioning *between* bumble bees and other flower visitors.

The strong explanatory power of tongue length with respect to selectivity in our study corroborates the established link between tongue length and floral choice. *I think it does so more strongly than most other field studyies I've read.* It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of selectivity on tongue length. Outliers worth talking about: telu-disc(+), muci-lip(-), muci-flag(+), soro/psit-head(+).

## Similarity of selectivity across species, potential for competition

While competition as a phenomenon arises from realized patterns of resource use, not the latent process of selection, insight into the process of selection may be a better predictor of competition. Use patterns alone are, in a sense, overfit to the observations at hand.

## Variability of floral abundance, visitation, and selectivity through elevation and time; its relevance to climate and land use change

Elevation-time patterns of visitation closely mirrored those of floral abundance, but the former were more diffuse; this may reflect the reluctance of bumble bees to switch from familiar floral taxa even when they become rarer [@Heinrich1976-vb] or less rewarding [I think I remember reading this somewhere...but there is also a paper showing that bumble bees switch more readily than honey bees].

The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula. Talk about idiosyncracies not capture dby morphology alone. importance of other traits, particularly color.

## Caveats:

The problem of queens vs. workers... Probably the best way to fend off this object would be to re-run all our analyses with only workers and see if any thing changes qualitatively. This could all go in the SI.

Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].

Inference of preference is sensitive to the scale at which availability is estimate [@Beyer2010-vr]. My floral survey data estimates availability more or less at the meadow scale, right? But I don't know what is available outside the meadow. This problem is also discussed in [@Harmon-Threatt2017-bz], specifically in the context of bees and flowers.

The sensitivity of "preference" to availability [@Johnson1980-co; @Beyer2010-vr].

*First, the estimates ofpreference  are  not  directly  interpretable  as  reflectingthe  intrinsic  behavioural  motivations  of  the  animal.Positive  (or  negative)  preference  for  a  given  habitatimplies only that the habitat is used more (or less) fre-quently than expected by chance,conditional upon thedefined sample of availability.  Even  high  levels  of  useof a habitat can be associated with negative coefficientsif that habitat is common, and, conversely, low levelsof  use  can  be  associated  with  positive  coefficientsif the habitat is rare. The sign and magnitude of thecoefficient   cannot   be   interpreted   with   respect   tothe   behavioural   decision   the   animal   is   making.* [@Beyer2010-vr].

*The second consequence is that estimates of prefer-ence  based  on  different  samples  of  availability  (e.g.Design III inManlyet al.2002) are not directly com-parable.  The  sign  and  magnitude  of  the  estimate  ofpreference  is  a  function  of  both  the  sample  of  useand the sample of availability. This creates an identifia-bility  problem  because  preference  can  change  as  afunction of changing use and/or changing availability.If   the   sample   of   availability   differs   between   twoestimates of preference, it is meaningless to draw infer-ences based on the difference in preference. Differentpreferences can arise from the same selection process,and   similar   preferences   can   arise   from   differentselection processes.* [@Beyer2010-vr].

This is exactly the problem I am facing. I think, though, I can present my "preference lability analysis" together with the raw visitation and availability data, discuss this limitation explicitly, and let the reader decide what to make of it. This is also discussed very clearly, and in the context of bumblebees, in @Harmon-Threatt2017-bz.

*Given that preference is conditional on avail-ability,  that  availability  is  likely  to vary  in  space  andthat preference is likely to change as a function of avail-ability, there is little theoretical basis for believing thatpreferences estimated in one area will be good predic-tors  of  preference  in  others.* [@Beyer2010-vr].

In answer to this, I can point to the largely agreeing results of other bumble bee studies.

*Models of habitat preference provide a phenomenolo-gical  description  of  habitat  use,  conditional  on  thesample of availability. Although some of the simplify-ing  assumptions  upon  which  these  models  are  often based  may  be  biologically  unrealistic,  they  can  stillhave predictive power (Boyceet al.2002;Wienset al.2008). The degree, however, to which simple modelsof habitat preference help us to develop a mechanisticunderstanding of animal movement and distribution islimited. Mechanistic models are required to advanceour  biological  understanding  of  the  behavioural  pro-cesses  that  result  in  animal  movement,  distributionand habitat use.*

Tongue length is my mechanistic model. And I think the observation that the mechanistic model better explains use-availability bias than raw selection indicates that the use-availability approach better captures the way bumble bees make foraging decisions.

Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].

*The problem of conflating preference and niche partitioning:*

Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest. This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn]. Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change.

Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr]. In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length.

@@

## What we did and why

In this study, we investigated patterns of floral resource selection in a mountain bumble bee community with the goals of (1) identifying the most selected floral morphotypes, (2) evaluating the role of floral abundance in driving floral selection, (3)

@@

We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite null modeling.

>>>>

Sampling of bumble- bee-wildflower interactions consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Sampling was only conducted during periods without rain and with an air temperature of at least 6°C. Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors. Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz. While male bumble bees were recorded, we chose to analyze only the pooled visitation of queens and workers.

In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro. While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012. The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July. In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient. See Appendix X for a full description of sampling patterns.

@@

We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of selectivity between bumble bee species and floral morphotypes. We used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values. Then, using the results of this analysis, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf]. To verify that our inference of selectivity was not sensitive to our decision to pool queen and worker visitation, we recalculated the null model and tongue length HGAM using only data from workers. Finally, we compared morphotype selectivity across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype selectivity matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.

@@

The omission of queens from the calculation of overall selectivity did not yield qualitatively different results, so we present only the results of the full data set. The results of the worker-only analysis are included in Appendix X.

@@

Again, these results did not differ qualitatively when selectivity was based on the worker-only null model, so we present only the results of the full data set. The results of the worker-only analysis are included in Appendix X.

@@

It is well-known that bumble bees forage non-randomly accordingly to morphological constraints, innate preferences, and experienced rewards [e.g. Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm]. In this regard, the results of our null model analysis are as unsurprising as they are unambiguous: bumble bee resource selection almost never represents flora in unaltered proportion to their abundance. The *target use hypothesis* [@De_Valpine2013-be; @Harmon-Threatt2017-bz], however, extends this idea to the perfect antithesis of the random foraging model, positing that bumble bee foraging is a process of *pure* preference, completely irrespective of resource abundance.

Given the well-established mechanistic link between bumble bee tongue length and floral selection [e.g. @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained selectivity (i.e. null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice. This casts doubt on the *target use hypothesis* as a general rule of bumble bee foraging. It does not, however, mean that the *target use hypothesis* does not obtain under certain circumstances, such as when a highly preferred resource varies within a nonlimiting rage of availability. Moreover, our results suggest something akin to the *target use hypothesis* but formulated negatively with respect to resources that are *not selected*. Visitation rates to disc flowers in our system, for example, were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not. In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to parse these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging. 
<!-- Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study. -->

## 4.2 Overall patterns of selectivity

We found that positive selection was concentrated on 3 floral morphotypes: bell, lip, and flag flowers. These results are broadly consistent with the raw visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg, Neumayer1999-il in the Austrian Alps ~55 km south of our study system, and Farnberger1996-xo in the vicinity of Vienna, suggesting that our findings are at least regionally generalizeable. These neighboring studies tended to find a stronger role of head flowers than suggested by our selectivity analysis, but this is probably because they characterized foraging in terms of raw visitation rather than selectivity. Indeed, our raw visitation data similarly indicate a stronger role of head flowers, but our selectivity analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.

When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera), and especially the hover flies of the family Syrphidae, are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g. disc flowers) that receive less attention from bumble bees [@Neumayer1999-il]. In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper flower morphotypes. The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.

While bell, lip, and flag flowers were the most selected morphotypes on average, selectivity toward these morphotypes varied broadly across bumble bee species. Moreover, while average selectivity toward funnel and head flowers was low, each received strongly positive selection from one or two bumble bee species. The majority of variation in selectivity toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes selected more positively by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes selected more positively by medium- and long-tongued bumble bee species. The failure of tongue length to explain selectivity toward disc and stalk-disc flowers reflects the fact that these flowers were almost unanimously ignored by all bumble bee species. We cannot say from our data why disc and stalk-disc flowers received such consistently low visitation, but the pattern may be driven in part by the fact that some disc flowers offer pollen but little or no nectar; indeed, Kugler's (1970) typology recognizes this distinction with subdivisions of the disc morphotype into those with (class 1.2) and without (class 1.1) nectar rewards. A limitation of our study is that we could not reliably distinguish pollen- and nectar-foraging bumble bees, and it is likely that this distinction would shed further light on the mechanisms underlying the selection patterns we observed.

The strong explanatory power of tongue length with respect to selectivity in our study corroborates the established link between tongue length and floral choice [e.g. @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn]. It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of selectivity on tongue length. *B. terrestris/lucorum* showed comparatively high (though still negative) selectivity toward disc flowers. *B. mucidus* exhibited unusually low selectivity (though not significantly different from neutrality) toward lip flowers, complemented by unusually high selectivity toward flag flowers. *B. soroeensis* and the *B. psythyrus* species-group were anomalous in their strongly positive selection toward head flowers. Such unexplained variation in selectivity may indicate that traits not captured by Kugler's (1970) morphotypology are also important in governing floral selection by bumble bees. Indeed, one such trait --- floral color --- is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq]. This effect can also be seen in field studies, where bumble bees have been shown to over- or under-represent certain floral colors [e.g. @Teras1976-pi; @Neumayer1999-il]. Moreover, since floral color can be correlated with floral morphology [@Teras1976-pi], it is possible that some of the selectivity attributed to morphology in our models could represent a latent effect of floral color.

As mountain bumble bees can be expected to face a future of declining resource availability and altered elevational distributions, it is relevant to ask which species seem poised for the greatest resource conflict should competitive conditions arise. While any answer to this question should be regarded as speculative, considering the similarities of morphotype selectivity between species is a good place to start. As a general rule, it can be said that species of similar tongue length will be, on average, more likely to compete for resources than species of dissimilar tongue length. Our ordination analysis, though, suggests that there may be significant patterns of similarity and divergence even among species of similar tongue length. Among short-tongued species, *B. jonellus* and *B. pyrenaeus* cluster tightly around bell flowers, while *B. psithryus* (species group) and *B. soroeensis* cluster tightly around head flowers. Medium- and long-tongued species differ in their allocation of selectivity across funnel, lip, and flag flowers. *B. gerstaeckeri*, for example, is associated with lip and funnel flowers but eschews flag flowers, while *B. mucidus* does this opposite. With respect to the question of climate-induced competition, the case of *B. gerstaeckeri* deserves special consideration. A red-listed vulnerable species, *B. gerstaeckeri* is a pollen specilaist on the lip flowers of the genus *Aconitum* [@Ponchau2006-af; @Schneller2014-ry], and there is evidence that joint elevational shifts of bumble bees and floral resources over the past century have generated more intense exploitation of *Aconitum*, which may indicate nascent competitive conditions for *B. gerstaeckeri* [@Marshall2020-pc].

Broaching the topic of competition raises an issue that can confound any observational study of resource selection: how to disentangle the endogenous process of resource preference from the exogenous process of niche partitioning driven by competition [@Pianka1974-qk]. Empirical selectivity could reflect a latent process of resource preference, as we assume in our interpretation. Alternatively, though, visitation patterns and the selectivity inferred from them could be distorted by competitive forces. Intraspecific competition in a bumble bee community would tend to force individuals of a species to broaden their diet beyond what they would prefer under non-competitive conditions, while interspecific competition would be expected to cause a narrowing of diet breadth toward resources that can be exploited most efficiently. Thus, observed visitation patterns, even when calculated as abundance-relative selectivity, can only be interpreted in terms of *preference* to the extent that competition can be assumed to be negligible, and assumption must be justified on the basis of additional knowledge of the study system under consideration. Our study system in the Berchtesgadener Alps is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn]. We cannot, however, exclude the possibility that competitive forces distort, to some degree, the visitation patterns we observed. Even when competitive conditions do not obtain universally in a given system, competition may occur in a spatially and/or temporally localized way [@Ranta1981-vx], perhaps associated with extreme climatic events that cause reductions in floral resource avaulability [@Rasmont2012-zq; @Thomson2016-gq; @Phillips2018-me]. This calls to mind the insight of MacArthur1972-qz, that competition need not be frequent to have lasting ecological impacts, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today.

## 4.3 Variability of floral abundance, visitation, and selectivity through elevation and time

*First, talk about the lability of selectivity, interesting difference between species, caveats.*

General patterns

While the floral choices of bumble bees have been studied extensively in both laboratory [e.g. @Kugler1943-uy; @Raine2007-nq; @Maharaj2019-wm] and field studies [e.g. @Heinrich1976-wc; @Teras1976-pi; @Goulson2008-zy], studies of the *dynamics* of bumble bee floral choice along ecological gradients are rare [e.g. @Miller-Struttmann2014-wz], and we are not aware of any that use null model controls to distinguish selectivity from the neutral effects of relative abundance.

Interesting deviations

The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.

We reiterate, however, the inherent problems of comparing selectivity --- understood as an index of the latent process of preference --- across different contexts of resource abundance. We recommend the cautious interpretation of selectivity patterns through elevation and time in conjunction with the underlying patterns of floral abundance and raw visitation.

*Then, segue to climate and land use.*

Biological patterns through elevation and time are acted upon by climate and land use change, and the consequence of this action is the generation of new biological patterns through elevation and time. A central task of ecology is to understand the patterns that will be acted upon and predict the ecological ramifications of the patterns that will be created. In this regard, it is significant that we find marked structure in the morphotypal composition. raw visitation, and selectivity of the floral community on which the bumble bees of the Berchtesgadener Alps depend. While our data permit only speculation as to the climate and land use driven processes at work in our system, our speculation can be nonetheless well-informed. Climate change over the last century has caused significant upslope movement [e.g. @Marshall2020-pc] and phenological shifts [e.g. @CaraDonna2014-nq] of floral resources for pollinators. It is likely that the current patterns of morphotype relative abundance that we observed differ from those that would have been observed a century ago, and in the coming decades warming climates will probably induce further shifts upslope [@Marshall2020-pc] and potentially complex shifts in the onset, peak, and end of flowering periods [@CaraDonna2014-nq]. These effects of climate change will interact with the likely more rapid effects of land use change. In particular, the reforestation of montane meadows can be expected to erode the abundance of flowers from mid-elevation up to the tree line (in our system, around ~1500 m.s.l.). Above the tree line, however, floral communities should be relatively safe from reforestation since the advance of alpine tree lines tends to lag behind climate change [@Dullinger2004-qu]. At lower elevations, agricultural intensification can be expected to suppress overall floral availability by favoring grasses over forbs and shift the composition of the community toward species that thrive under the altered soil conditions and disturbance regimes of intensive agriculture [@Tasser2002-om].

## 4.4 Final caveats

In addition to the qualifying remarks provided in the sections above, it is important to acknowledge some limitations of our study that should caution the reader against the over-interpretation of our results. First, the inference of selection as use-relative-to-abundance is sensitive to exactly *how* abundance is estimated [@Johnson1980-co; @Beyer2010-vr]. Of particular concern in our study is the problem of spatial scale. By necessity, we measured floral abundance and visitation along transects *within* our study sites. Bumble bee foraging range, however, is much larger than our study sites, and it is possible that the relative abundances of floral morphotypes outside our study sites but within foraging range differ from our within-site estimates. Moreover, because bumble bees can forage across multiple meadow patches in a mountain forest matrix [@Mola2020-yc], there exists a prior selective process --- Johnson's (1980) "second-order selection" --- in which bumble bees choose among available foraging areas before selecting from resources within an area. Finally, to account for the neutral effects of abundance on floral use, we used a simple null model in which the probability of a resource being selected was assumed to be linearly proportional to the relative abundance of the resource. It is possible, however, for resource use to exhibit nonlinear "functional responses" @Mysterud1998-rs to resource abundance, which, relative to a linear null model, would give biased estimates of selectivity. Indeed, the fact that patterns of selectivity through elevation and time **(Figure 8)** partly retain of the pattern of resource abundance **(Figure 7)** suggests that there may be some nonlinear dependence of usage on abundance that cannot fully be removed by our linear null model.

@@

Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations. Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.

<!-- Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr]. In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length. -->

@@

We thank the National Park Berchtesgaden for their support and the permission to work in the national park, and we thank the owners of the study sites outside the national park for access to their land. The field work of this study was performed within the framework of the joint research center FORKAST as part of the ‘Bavarian Climate Programme 2020'. D.J. McNeil provided for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite null modeling.
